---
title: "data_vis"
author: "Jerri Chen"
date: "11/12/19"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(plotly)
library(viridis)
library(tableone)
library(formattable)
library(data.table)

customGreen0 = "#DeF7E9"

customGreen = "#71CA97"

customRed = "#ff7f7f"
```

## Data import 

```{r, message=FALSE, warning=FALSE}
nsqipspineCP_1617 = read_csv("./data/nsqipspineCP_1617.csv")
```


## Data simplification and coding

```{r, message=FALSE, warning=FALSE}
cp_spine_tidy = nsqipspineCP_1617 %>%
  mutate_if(is.numeric, ~replace(., . == -99, NA)) %>%
  mutate(
    age_years = age_days/365.25,
    height = height*2.54,
    weight = weight/2.2) %>%
  mutate(
    bmi = weight/((height/100)*(height/100)),
    asa_status = case_when(
      asaclas == "ASA 1 - No Disturb" ~ "1",
      asaclas == "ASA 2 - Mild Disturb" ~ "2",
      asaclas == "ASA 3 - Severe Disturb" ~ "3",
      asaclas == "ASA 4 - Life Threat" ~ "4",
      asaclas == "None assigned" ~ "NA"),
    home_discharge = case_when(
      dischdest == "Expired" ~ "FALSE",
      dischdest == "Facility Which was Home" ~ "TRUE",
      dischdest == "Home" ~ "TRUE",
      dischdest == "Rehab" ~ "FALSE",
      dischdest == "Separate Acute Care" ~ "FALSE",
      dischdest == "Skilled Care, Not Home" ~ "FALSE",
      dischdest == "Unknown" ~ "NA",
      dischdest == "Unskilled Facility Not Home" ~ "FALSE",
      dischdest == "NULL" ~ "NA"),
    level_13 = case_when(
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 2 TO 3 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 4 TO 7 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 8 OR MORE VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; UP TO 6 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 7 TO 12 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 13 OR MORE VERTEBRAL SEGMENTS" ~ "TRUE")) %>%
  filter(home_discharge != "NA") %>% 
  select(pufyear_x:ped_spn_post_neurodeftype, age_years, sex, height, weight, bmi, ethnicity_hispanic, race, asa_status, transt, ventilat, asthma, hxcld, oxygen_sup, crf, impcogstat, seizure, nutr_support, hemodisorder, level_13, optime, tothlos, d_opto_dis, death30yn, supinfec, wndinfd, orgspcssi, dehis, oupneumo, pulembol, renainsf, urninfec, cszre, neurodef, cdarrest, othbleed, bleed_ml_tot, othcdiff, othsysep, unplannedreadmission1, reoperation, dischdest, home_discharge)
```


## Patient Demographics
### For Visualization

Count of patients who did not have home discharge.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>%
  summarize(n = n()) %>% 
  plot_ly(
    x = ~home_discharge, y = ~n, type = "bar"
    )
```

Stratified by sex.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(sex) %>% 
  plot_ly(
    x = ~home_discharge, y = ~n, color = ~sex, type = "bar"
    )
```

Age mean and SD of patients who did and did not have home discharge.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(sex, home_discharge) %>% 
  summarize(mean_age = mean(age_years), sd = sd(age_years)) %>% 
  plot_ly(
    x = ~home_discharge, y = ~mean_age, color = ~sex, type = "bar"
    )
```

Distribution of age in patients who did and did not have home discharge.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(y = ~age_years, color = ~home_discharge, type = "box",
          colors = "Set2")
```

Weight mean and SD of patients who did and did not have home discharge.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(sex, home_discharge) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE), sd = sd(weight, na.rm = TRUE))
```

Race distribution.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(race) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names() %>% 
  plot_ly(
    x = ~race, y = ~n, color = ~home_discharge, type = "bar"
    )
```

Playing around with how to represent this data. Here is a plot of just the distribution of race in the dataset, without stratification.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  count(race) %>% 
  plot_ly(
    x = ~race, y = ~n, type = "bar"
    )
```
And then this shows the percent of patients in each race group who and did not have home discharge. We could show these side by side? Percent is probably better than raw numbers since that is harder to interpret accross groups.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(race) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_nothome = (false/(true + false))*100) %>% 
  replace(is.na(.), 0) %>% 
  plot_ly(
    x = ~race, y = ~percent_nothome, type = "bar"
    )
```


Ethnicity distribution stratified by home vs. not home discharge. 
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(ethnicity_hispanic) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))  %>% 
  plot_ly(
    x = ~ethnicity, y = ~n, color = ~home_discharge, type = "bar"
    )
```

Similar to above, we could also choose to just make two plots. Maybe that is clearer?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  count(ethnicity_hispanic) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))  %>% 
  plot_ly(
    x = ~ethnicity, y = ~n, type = "bar"
    )
```

```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(ethnicity_hispanic) %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  )) %>% 
    pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_nothome = (false/(true + false))*100) %>% 
  replace(is.na(.), 0) %>% 
  plot_ly(
    x = ~ethnicity, y = ~percent_nothome, type = "bar"
    )
```



Diagnosis distribution.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(ped_spn_class) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names() %>% 
  plot_ly(
    x = ~ped_spn_class, y = ~n, color = ~home_discharge, type = "bar"
    )
```

Association between operation time and bleeding?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~optime, y = ~bleed_ml_tot, type = "scatter", mode = "markers",
    color = ~home_discharge, alpha = 0.5)
```

Association between bleeding and length of stay?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~bleed_ml_tot, y = ~tothlos, type = "scatter", mode = "markers",
    color = ~home_discharge, alpha = 0.5)
```

Association between age and operation time?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~age_years, y = ~optime, type = "scatter", mode = "markers",
    color = ~home_discharge, alpha = 0.5)
```

### For Tables (Pivot Wider)
```{r, message=FALSE, warning=FALSE}
data_discharge = cp_spine_tidy %>%
  group_by(home_discharge) %>%
  summarize(n = n()) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names()%>% 
  mutate(percent = (false/(true + false))*100)
```

```{r, message=FALSE, warning=FALSE}
data_sex = cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(sex) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent = (false/(true + false))*100)
```

```{r, message=FALSE, warning=FALSE}
data_age = cp_spine_tidy %>%
  group_by(sex, home_discharge) %>% 
  summarize(mean_age = mean(age_years), sd = sd(age_years))
```

```{r, message=FALSE, warning=FALSE}
data_weight = cp_spine_tidy %>%
  group_by(sex, home_discharge) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE), sd = sd(weight, na.rm = TRUE))
```

```{r, message=FALSE, warning=FALSE}
data_race = cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(race) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_nothome = (false/(true + false))*100) %>% 
  replace(is.na(.), 0)
```

```{r, message=FALSE, warning=FALSE}
data_ethnicity = cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(ethnicity_hispanic) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))
```
 
```{r, message=FALSE, warning=FALSE}
data_diagnosis = cp_spine_tidy %>%
  group_by(home_discharge) %>% 
  count(ped_spn_class) %>% 
  pivot_wider(
    names_from = "home_discharge",
    values_from = "n"
  ) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names()
```
 
 
 Create clean data frame for table 1, create table
```{r}
cp_spine_table1 = cp_spine_tidy %>%
mutate(
    sex = factor(sex, ordered = TRUE, levels = c("Female", "Male")),
    race = factor(race, ordered = FALSE, levels = c("American Indian or Alaska Native", "Asian", "Black or African American", "Native Hawaiian or Other Pacific Islander", "Unknown/Not Reported", "White")),
    admit_from = factor(transt, ordered = TRUE, levels = c("Admitted from home/clinic/doctor's office", "Admitted through ER including outside ER with direct hospital admission", "Chronic care/Rehab/Intermediate Care/Spinal Cord", "Transferred from outside hospital (NICU, PICU, Inpatient on General floor, Adult)", "Other")),
    ASAstatus = factor(asa_status, ordered = TRUE, levels = c("1", "2", "3", "4", "5")),
    ventilator_dependence = factor(ventilat, ordered = TRUE, levels = c("No", "Yes")),
    asthma = factor(asthma, ordered = TRUE, levels = c("No", "Yes")),
    home_oxygen = factor(oxygen_sup, ordered = TRUE, levels = c("No", "Yes")),
    cognitive_impairment = factor(impcogstat, ordered = TRUE, levels = c("No", "Yes")),
    seizure_disorder = factor(seizure, ordered = TRUE, levels = c("No", "Yes")),
    nutritional_support = factor(nutr_support, ordered = TRUE, levels = c("No", "Yes")),
    hematologic_disorder = factor(hemodisorder, ordered = TRUE, levels = c("No", "Yes"))
)

myVars <- c("age_years", "sex", "height", "weight", "ASAstatus", "admit_from", "ventilator_dependence", "asthma", "home_oxygen", "cognitive_impairment", "seizure_disorder", "nutritional_support", "hematologic_disorder")
catVars <- c("sex", "ASAstatus", "admit_from", "ventilator_dependence", "asthma", "home_oxygen", "cognitive_impairment", "seizure_disorder", "nutritional_support", "hematologic_disorder")

tab3 <- CreateTableOne(vars = myVars, data = cp_spine_table1, factorVars = catVars)

print(tab3)
```