---
title: "Data_Analysis_CPSPINE"
Author: "Katherine Dimitropoulou"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(viridis)
```


## Data import 

```{r, message=FALSE, warning=FALSE}
nsqipspineCP_1617 = read_csv("./nsqipspineCP_1617.csv")
```

## Initial Data Cleaning & Selection of a subset of data to examine the purpose of the study 

```{r, message=FALSE, warning=FALSE}
cp_spine_tidy = nsqipspineCP_1617 %>%
  mutate_if(is.numeric, ~replace(., . == -99, NA)) %>%
  mutate(
    age_years = age_days/365.25,
    height = height*2.54,
    weight = weight/2.2) %>%
  mutate(
    bmi = weight/((height/100)*(height/100)),
    asa_status = case_when(
      asaclas == "ASA 1 - No Disturb" ~ "1",
      asaclas == "ASA 2 - Mild Disturb" ~ "2",
      asaclas == "ASA 3 - Severe Disturb" ~ "3",
      asaclas == "ASA 4 - Life Threat" ~ "4",
      asaclas == "None assigned" ~ "NA"),
    home_discharge = case_when(
      dischdest == "Expired" ~ "FALSE",
      dischdest == "Facility Which was Home" ~ "TRUE",
      dischdest == "Home" ~ "TRUE",
      dischdest == "Rehab" ~ "FALSE",
      dischdest == "Separate Acute Care" ~ "FALSE",
      dischdest == "Skilled Care, Not Home" ~ "FALSE",
      dischdest == "Unknown" ~ "NA",
      dischdest == "Unskilled Facility Not Home" ~ "FALSE",
      dischdest == "NULL" ~ "NA"),
    level_13 = case_when(
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 2 TO 3 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 4 TO 7 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 8 OR MORE VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; UP TO 6 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 7 TO 12 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 13 OR MORE VERTEBRAL SEGMENTS" ~ "TRUE"))  %>% 
  filter(home_discharge != "NA") %>% 
  select(pufyear_x:ped_spn_post_neurodeftype, age_years, sex, height, weight, bmi, ethnicity_hispanic, race, asa_status, transt, ventilat, asthma, hxcld, oxygen_sup, crf, impcogstat, seizure, nutr_support, hemodisorder, level_13, optime, tothlos, d_opto_dis, death30yn, supinfec, wndinfd, orgspcssi, dehis, oupneumo, pulembol, renainsf, urninfec, cszre, neurodef, cdarrest, othbleed, bleed_ml_tot, othcdiff, othsysep, unplannedreadmission1, reoperation, dischdest, home_discharge)
```
##Analysis and models

Our goal is to develop a model,in which we can use important pre-operative factors to predict the patient's discharge at home. 
Selection of predictors to include in the initial model
Data processing, elimination of missing data, and data selection for the model. We observe that all cp patients in the dataset were patient who came from home. 

```{r, message=FALSE, warning=FALSE}
model_data = cp_spine_tidy %>% 
  mutate(
    sex1=case_when(
    sex=="Male"~"0",
    sex=="Female"~"1"),
    sex1=as.factor(sex1),
    race = case_when(
    race == "American Indian,Alaskan Native" ~ "1",
    race == "Asian" ~ "2",
    race == "Black or African American" ~ "3",
    race == "Native Hawaiian or Other Pacific Islander" ~ "4",
    race == "White" ~ "5",
    race == "Unknown Not Reported" ~ "6"),
    race1=as.factor(race),
    asa_level = ifelse(asa_status >2, 1, 0),
    asa_level=as.factor(asa_level),
    ventilat1=case_when(
    ventilat=="No"~"0",
    ventilat=="Yes"~"1"),
    ventilat1=as.factor(ventilat1), 
  asthma1=case_when(
    asthma=="No"~"0",
    asthma=="Yes"~"1"),
  asthma1=as.factor(asthma1),
  hxcld1=case_when(
    hxcld=="No"~"0",
    hxcld=="Yes"~"1"),
  hxcld1=as.factor(hxcld1),
  oxygen_sup1=case_when(
    oxygen_sup=="No"~"0",
    oxygen_sup=="Yes"~"1"),
  oxygen_sup1=as.factor(oxygen_sup1),
  seizure1=case_when(
    seizure=="No"~"0",
    seizure=="Yes"~"1"),
  seizure1=as.factor(seizure1),
  nutr_support1=case_when(
    nutr_support=="No"~"0",
    nutr_support=="Yes"~"1"),
  nutr_support1=as.factor(nutr_support1),
hemodisorder1=case_when(
    hemodisorder=="No"~"0",
    hemodisorder=="Yes"~"1"),
  hemodisorder1=as.factor(hemodisorder1),
 level_13=as.factor(level_13),
 home_discharge1=case_when(
   home_discharge=="TRUE"~"1",
   home_discharge=="FALSE"~"0"),
home_discharge1=as.factor(home_discharge1), 
  crf1 = case_when(
      crf == "Major cardiac risk factors" ~ "FALSE",
      crf == "Severe cardiac risk factors" ~"FALSE",
      crf == "Minor cardiac risk factors" ~ "TRUE",
      crf == "No cardiac risk factors" ~ "TRUE"), 
  crf1=as.factor(crf1),
  transt1 = case_when(
      transt == "Chronic care Rehab/Intermediate Care/Spinal Cord" ~ "FALSE",
      transt == "Transferred from outside hospital (NICU, PICU, Inpatient on General floor, Adult" ~"FALSE",
      transt == "other" ~ "FALSE",
      transt == "Admitted from home/clinic/doctor's office" ~ "TRUE",
      transt == "Admitted through ER, including outside ER with direct hospital admission" ~ "TRUE"),
  transt1=as.factor(transt1))%>%
  select(age_years, bmi, weight, sex1, transt1, crf1, race1, asa_level, ventilat1, asthma1, hxcld1, oxygen_sup1, seizure1, nutr_support1, hemodisorder1, level_13, home_discharge1) 
```
Overall quick descriptives for the data entered in the model suggest that all cp patients came from home.
```{r, message=FALSE, warning=FALSE}
summary(model_data)
```
##NEED TO ENTER HERE A GRAPH OF THE ALL PREDICTORS WITH THE OUTCOME

Initially there were 19 pre-operative variables that could predict the main outcome `home_discharge` To select the variables for our model we used findings from our literature review and results from preliminary analysis (crosstabs). From groups of variables that were highly correlated with each other, we  selected one variable that was supported from the literature. For example, we used weight instead of height and bmi, as height is a non reliable measure for individuals with CP and has a lot of missing values in the dataset. After the cross tabs exploratory analysis (see CPspine extensive), we eliminated variables that presented with 0 data in at least one of the cells. We combined multilevel variables to binary variables (i.e. crf) due to small sample size. We finally eliminated variables such as `ethnicity` because there was a large number of patients who self-identified as "Other" and it is unclear how to use "Other" to create preditions. Therefore, our full regression model consisted of 12 variable. 
##Logistic Regression Models
Full Model
```{r, message=FALSE, warning=FALSE}
mylogit <- glm(home_discharge1 ~ age_years+ weight+ sex1 + crf1+ventilat1 + asthma1 + hxcld1 + oxygen_sup1 + seizure1 + nutr_support1 + hemodisorder1+ level_13, data = model_data, family = "binomial")
summary(mylogit)
```
We further examined ways to improve our model fit, and eliminated variables that had cell counts 
 <=5) and those that had no support from the literature to be predictors for this outcome. The reduced model included 3 predictors.
 
Reduced Model
```{r, message=FALSE, warning=FALSE}
mylogit <- glm(home_discharge1 ~ age_years+ weight+hemodisorder1, data = model_data, family = "binomial")
summary(mylogit)
```
We tested which of the two models (Full or Reduced) fits better. We see the deviance residuals, which are a measure of model fit. This part of output shows the distribution of the deviance residuals for individual cases used in the model. Compared to the other two models this fits better. AIC is also better (lower). Age variable is significant at a=.01,  the pre-existance of a hematologic disorder variable is significant at a=.01.
In addition, we wanted to see if our model fits the data significantly better than a model with just the intercept (a null model). 

We use the null and deviance residuals 
```{r, message=FALSE, warning=FALSE}
with(mylogit, null.deviance - deviance)
```
with Df
```{r, message=FALSE, warning=FALSE}
with(mylogit, df.null - df.residual)
```
We  compute the Chi-Square test
```{r, message=FALSE, warning=FALSE}
with(mylogit, pchisq(null.deviance - deviance, df.null - df.residual, lower.tail = FALSE))
```
The chi-square provides a p-value=0.000367 (p< 0.001), thus, our model as a whole fits significantly better than the null model. 

We can use the likelihood ratio test (-2*log likelihood) to see the model’s log likelihood:
```{r, message=FALSE, warning=FALSE}
logLik(mylogit)
```
##Preliminary intrepretation
1. For every 1 year increase in age the log odds of homedischarge (versus discharge to other healthcare facilities) decreases by 0.19.
2. Having a pre-existing hematologic disorder decreases the log odds of being discharged to home by 1.275.
Confidence intervals for the log odds coeficients (log-likelihood function)
```{r, message=FALSE, warning=FALSE}
confint(mylogit)
```
Confidence intervals using the standard errors
```{r, message=FALSE, warning=FALSE}
confint.default(mylogit)
```
Exponantiate to odds ratios and Confidence intervals
```{r, message=FALSE, warning=FALSE}
exp(cbind(OR = coef(mylogit), confint(mylogit)))
```
##Data interpretation
For every one year  increase in age, the odds of being discharged home (versus being discharged to another health care facility) increase by 0.83. This means that for every year increase in age the odds of being release to a health care facility (not home) increase by (1/.83=1.2) by 20%.
The odds of being discharged at home if you have a hemodisorder are 0.279.This means that if the patient has a a pre-existing hematologic disorder the odds of being discharged to a health care facility are (1/.279=3.58) 3.58 the odds of being discharged to home.

