---
title: "data_vis"
author: "Jerri Chen"
date: "11/12/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(plotly)
library(viridis)
```

## Data import 

```{r, message=FALSE, warning=FALSE}
nsqipspineCP_1617 = read_csv("./data/nsqipspineCP_1617.csv")
```

note: Where variable type and length disagree, we get error message. Coercion occurs and outcome is all numeric/all character for any given column with longest length.

## Data simplification and coding

```{r, message=FALSE, warning=FALSE}
cp_spine_tidy = nsqipspineCP_1617 %>%
  mutate_if(is.numeric, ~replace(., . == -99, NA)) %>%
  mutate(
    age_years = age_days/365.25,
    height = height*2.54,
    weight = weight/2.2) %>%
  mutate(
    bmi = weight/((height/100)*(height/100)),
    asa_status = case_when(
      asaclas == "ASA 1 - No Disturb" ~ "1",
      asaclas == "ASA 2 - Mild Disturb" ~ "2",
      asaclas == "ASA 3 - Severe Disturb" ~ "3",
      asaclas == "ASA 4 - Life Threat" ~ "4",
      asaclas == "None assigned" ~ "NA"),
    home_discharge = case_when(
      dischdest == "Expired" ~ "FALSE",
      dischdest == "Facility Which was Home" ~ "TRUE",
      dischdest == "Home" ~ "TRUE",
      dischdest == "Rehab" ~ "FALSE",
      dischdest == "Separate Acute Care" ~ "FALSE",
      dischdest == "Skilled Care, Not Home" ~ "FALSE",
      dischdest == "Unknown" ~ "NA",
      dischdest == "Unskilled Facility Not Home" ~ "FALSE",
      dischdest == "NULL" ~ "NA"
    )) %>%
  select(pufyear_x:ped_spn_post_neurodeftype, age_years, sex, height, weight, bmi, ethnicity_hispanic, race, asa_status, transt, ventilat, asthma, hxcld, oxygen_sup, crf, impcogstat, seizure, nutr_support, hemodisorder, optime, tothlos, d_opto_dis, death30yn, supinfec, wndinfd, orgspcssi, dehis, oupneumo, pulembol, renainsf, urninfec, cszre, neurodef, cdarrest, othbleed, bleed_ml_tot, othcdiff, othsysep, unplannedreadmission1, reoperation, dischdest, home_discharge)
```

## Patient Demographics
### For Visualization

Count of patients who needed re-operation within 30 days.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>%
  summarize(n = n()) %>% 
  plot_ly(
    x = ~reoperation, y = ~n, type = "bar"
    )
```

Stratified by sex.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(sex) %>% 
  plot_ly(
    x = ~reoperation, y = ~n, color = ~sex, type = "bar"
    )
```

Age mean and SD of patients who did and did not require reoperation.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(sex, reoperation) %>% 
  summarize(mean_age = mean(age_years), sd = sd(age_years)) %>% 
  plot_ly(
    x = ~reoperation, y = ~mean_age, color = ~sex, type = "bar"
    )
```

Distribution of age in patients who did and did not require reoperation.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(y = ~age_years, color = ~reoperation, type = "box",
          colors = "Set2")
```

Weight mean and SD of patients who did and did not require reoperation.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(sex, reoperation) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE), sd = sd(weight, na.rm = TRUE))
```

Race distribution.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(race) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names() %>% 
  plot_ly(
    x = ~race, y = ~n, color = ~reoperation, type = "bar"
    )
```

Playing around with how to represent this data. Here is a plot of just the distribution of race in the dataset, without stratification.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  count(race) %>% 
  plot_ly(
    x = ~race, y = ~n, type = "bar"
    )
```
And then this shows the percent of patients in each race group who required reoperation. We could show these side by side? Percent is probably better than raw numbers since that is harder to interpret accross groups.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(race) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_reop = (yes/(no + yes))*100) %>% 
  replace(is.na(.), 0) %>% 
  plot_ly(
    x = ~race, y = ~percent_reop, type = "bar"
    )
```

Ethnicity distribution stratified by reop vs no reop. 
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(ethnicity_hispanic) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))  %>% 
  plot_ly(
    x = ~ethnicity, y = ~n, color = ~reoperation, type = "bar"
    )
```

Similar to above, we could also choose to just make two plots. Maybe that is clearer?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  count(ethnicity_hispanic) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))  %>% 
  plot_ly(
    x = ~ethnicity, y = ~n, type = "bar"
    )
```

```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(ethnicity_hispanic) %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  )) %>% 
    pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_reop = (yes/(no + yes))*100) %>% 
  replace(is.na(.), 0) %>% 
  plot_ly(
    x = ~ethnicity, y = ~percent_reop, type = "bar"
    )
```



Diagnosis distribution.
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(ped_spn_class) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names() %>% 
  plot_ly(
    x = ~ped_spn_class, y = ~n, color = ~reoperation, type = "bar"
    )
```

Association between operation time and bleeding?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~optime, y = ~bleed_ml_tot, type = "scatter", mode = "markers",
    color = ~reoperation, alpha = 0.5)
```

Association between bleeding and length of stay?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~bleed_ml_tot, y = ~tothlos, type = "scatter", mode = "markers",
    color = ~reoperation, alpha = 0.5)
```

Association between age and operation time?
```{r, message=FALSE, warning=FALSE}
cp_spine_tidy %>%
  plot_ly(
    x = ~age_years, y = ~optime, type = "scatter", mode = "markers",
    color = ~reoperation, alpha = 0.5)
```

### For Tables (Pivot Wider)
```{r, message=FALSE, warning=FALSE}
data_reop = cp_spine_tidy %>%
  group_by(reoperation) %>%
  summarize(n = n()) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names()%>% 
  mutate(percent = (yes/(no + yes))*100)
```

```{r, message=FALSE, warning=FALSE}
data_sex = cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(sex) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent = (yes/(no + yes))*100)
```

```{r, message=FALSE, warning=FALSE}
data_age = cp_spine_tidy %>%
  group_by(sex, reoperation) %>% 
  summarize(mean_age = mean(age_years), sd = sd(age_years))
```

```{r, message=FALSE, warning=FALSE}
data_weight = cp_spine_tidy %>%
  group_by(sex, reoperation) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE), sd = sd(weight, na.rm = TRUE))
```

```{r, message=FALSE, warning=FALSE}
data_race = cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(race) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  mutate(percent_reop = (yes/(no + yes))*100) %>% 
  replace(is.na(.), 0)
```

```{r, message=FALSE, warning=FALSE}
data_ethnicity = cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(ethnicity_hispanic) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  janitor::clean_names() %>% 
  rename(ethnicity = ethnicity_hispanic) %>% 
  mutate(ethnicity = case_when(
    ethnicity == "No" ~ "Non-Hispanic",
    ethnicity == "Yes" ~ "Hispanic",
    ethnicity == "NULL" ~ "Other/Did Not Answer"
  ))
```
 
```{r, message=FALSE, warning=FALSE}
data_diagnosis = cp_spine_tidy %>%
  group_by(reoperation) %>% 
  count(ped_spn_class) %>% 
  pivot_wider(
    names_from = "reoperation",
    values_from = "n"
  ) %>% 
  replace(is.na(.), 0) %>% 
  janitor::clean_names()
```
 
## Code for Home Discharge

```{r, message=FALSE, warning=FALSE}
data_discharge = cp_spine_tidy %>%
  group_by(home_discharge) %>%
  summarize(n = n())
```
